---
title: "UK Fisheries Activity Explorer"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
runtime: shiny_prerendered
---

<style type="text/css">

.sidebar {  /* sidebar  */
   font-size: 13px;
   font-family: Arial;
}

</style>

<style>

.navbar {
  background-color:tan;
  border-color:mintcream;
}
.navbar-brand {
color:darkgreen!important;
}

</style>

<style>
.tabset {  /* tabset  */
   font-size: 13px;
   font-family: Arial;}
   

.page { /* page */

   font-size: 13px;
   font-family: Arial;
}
 </style> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
```{r, context="setup", message=FALSE, warning=FALSE, include=FALSE}
library(shiny)
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(plotly)
library(lubridate)
library(leaflet)
library(htmltools)
```

```{r, context="data", cache=TRUE, cache.extra=file.info("www/fisheriesData")$mtime, message=FALSE, warning=FALSE, include=FALSE}
# load FisheriesData date set to be used for the App

fisheriesData <- read.csv("www/fisheriesData.csv",header = TRUE,stringsAsFactors = FALSE)

fisheriesData$Date <- as.Date(fisheriesData$Date)

# create data frame for bubble graph over 12 months

fisheriesmonth <- mutate(fisheriesData, Month.number = month(fisheriesData$Date))

fisheriesym <- mutate(fisheriesmonth, Year = factor(year(fisheriesmonth$Date),levels= c("2014","2015","2016","2017")))

fisheries_trim <- fisheriesym %>% mutate(Month.name=month.abb[Month.number]) %>% group_by(Year, Month.number, Month.name) %>% summarise(Live.Weight=sum(Live.weight..tonnes./1000),Value=sum(Value..000s.))

fisheries_price <- mutate(fisheries_trim, Price = Value/Live.Weight)

# Landings into the UK by UK vessels 

subset_UK <- subset(fisheriesData, (Port.Nationality %in% c("England","Wales", "Scotland", "Northern Ireland")) & Vessel.Nationality == "UK")  

n1 <- subset_UK %>% tally() 

UKVesselsHome <-  subset_UK %>% mutate(Type = factor( rep("UK.vessels.fishing.in.the.UK",n1),levels="UK.vessels.fishing.in.the.UK"))

# Landings into the UK by foreign vessels  

subset_foreignVessels <- subset(fisheriesData, (Port.Nationality %in% c("England","Wales", "Scotland", "Northern Ireland")) & Vessel.Nationality != "UK")

n2 <- subset_foreignVessels %>% tally() 

foreignVessels <-  subset_foreignVessels %>% mutate(Type = factor(rep("Foreign.vessels.fishing.in.the.UK",n2),levels="Foreign.vessels.fishing.in.the.UK"))

# Landings abroad by UK vessels (Fishing abroad)

subset_UKabroad <- subset(fisheriesData, !(Port.Nationality %in% c("England","Wales", "Scotland", "Northern Ireland")) & Vessel.Nationality == "UK")

n3 <-subset_UKabroad %>% tally() 

UKVesselsAbroad <- subset_UKabroad %>% mutate(Type = factor(rep("UK.vessels.fishing.abroad",n3),levels="UK.vessels.fishing.abroad"))

fisheriesData1 <- rbind(UKVesselsHome,UKVesselsAbroad,foreignVessels) %>% mutate(Year = factor(year(as.Date(Date)),levels=c("2014","2015","2016","2017")))

fisheries_plot2 <- fisheriesData1  %>% group_by(Year, Type, Species.Group) %>% summarise(Live.Weight=sum(Live.weight..tonnes.),Value=sum(Value..000s.))

fisheries_priceplot2  <- fisheries_plot2  %>% mutate(Price=Value*1000/Live.Weight)


# Top 3 Species landed by the UK vessels in the UK and abroad
# Remove rows that contain 0. 

Top3Species.w <-  subset(subset(fisheriesData1, Type!="Foreign.Vessels.fishing.in.the.UK"), Live.weight..tonnes.!=0) %>%
group_by(Year, Species.Group, Species) %>% tally(Live.weight..tonnes.) %>% top_n(3) %>% rename(Live.Weight = n)

Top3Species.v <- subset(subset(fisheriesData1, Type!="Foreign.Vessels.fishing.in.the.UK"), Value..000s.!=0)  %>%
group_by(Year, Species.Group, Species) %>% tally(Value..000s.) %>% top_n(3) %>% rename(Value = n)

Top3Species <- subset(subset(fisheriesData1, Type!="Foreign.Vessels.fishing.in.the.UK"), Live.weight..tonnes.!=0)  %>%
group_by(Year, Species.Group, Species) %>% summarise (Live.Weight=sum(Live.weight..tonnes.),Value=sum(Value..000s.)) 

Top3Species.p <- Top3Species %>% 
group_by(Year, Species.Group, Species)  %>%  mutate(Price=Value*1000/Live.Weight) %>% tally(Price) %>% top_n(3) %>% rename(Price = n)

Top3Species_plot3 <- rbind(Top3Species.w,Top3Species.v,Top3Species.p)
```

Charts
===================================== 

Column {.sidebar}
-------------------------------------

This interactive dashboard allows the exploration of fisheries activity for UK commercial vessels landing into the UK and abroad, as well as fisheries activity for foreign commercial vessels landing into the UK.

Key statistics such as quantity (based on Live Weight) of fish landed, total worth, and average price per tonne are visualised. Live weight is the mass or weight of a product, when removed from the water.  

```{r, context = "render", echo=FALSE}

radioButtons("Criteria", h4("Choose criteria:"), choiceNames =c("Live Weight ('000 tonnes)", "Total Value (Million GBP)", "Price (GBP/tonne)"), choiceValues = c("Live Weight ('000 tonnes)", "Total Value (Million GBP)", "Price (GBP/tonne)"))

```

<br/> <br/>
<b>Author</b>: Ashley You <br/>
<b>Date</b>: 8th September 2017

Column {.tabset}
-------------------------------------

```{r, context ="server"}

# Fishing activity over month of the year  
string <- reactive(substring(input$Criteria,1,5))

output$plot1 <- renderPlotly ({ 
        if (string() == "Live ") { 
   p1 <-   ggplot(fisheries_trim, aes(x = Month.number, y = Live.Weight, fill = Year, colour = Year)) + geom_point(size = 6, shape = 18) + scale_x_continuous(name="Month of the Year", breaks=seq(1,12,1),labels= c("Jan", "Feb", "Mar", "Apr", "May", "Jun","Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) + scale_y_continuous(name="Live Weight in '000 tonnes") + ggtitle("Fisheries activity over 12 months (Live Weight)") + theme(plot.title = element_text(size=12, face="bold")) 
   
   ggplotly(p1)%>%layout(margin=list(l=70, b=120))
                }
        
        else if (string() == "Total") {
    p1 <-   ggplot(fisheries_trim, aes(x = Month.number, y = Value/1000, fill = Year, colour = Year)) + geom_point(size = 6, shape = 21, colour = "mediumvioletred") + scale_x_continuous(name="Month of the Year", breaks=seq(1,12,1),labels= c("Jan", "Feb", "Mar", "Apr", "May", "Jun","Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) + scale_y_continuous(name="Total Fish Value in Million GBP") + ggtitle("Fisheries activity over 12 months (Total Value)") + theme(plot.title = element_text(size=12, face="bold"))    

   ggplotly(p1)%>%layout(margin=list(l=70, b=120))
        }
        
        else {
p1 <- ggplot(fisheries_price, aes(x = Month.number, y = Price, fill = Year, colour = Year)) + geom_point(size = 6, shape = 18) + scale_x_continuous(name="Month of the Year", breaks=seq(1,12,1),labels= c("Jan", "Feb", "Mar", "Apr", "May", "Jun","Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) + scale_y_continuous(name="Average Fish Price (GBP per tonne)") + ggtitle("Average fish price over 12 months") + theme(plot.title = element_text(size=12, face="bold")) 
                
 ggplotly(p1)%>%layout(margin=list(l=70, b=120))
        }
})

# Comparing fishing activity of UK vessels in the UK and abroad, and foreign vessels in the UK 

output$plot2 <- renderPlotly({
                
                if (string() == "Live ") { 
        p2 <- ggplot(fisheries_priceplot2,aes(x=Year,y = Live.Weight/1000, fill=Type)) + geom_bar(position="dodge",stat="identity") + ggtitle("Landings over the years by species group") + labs(x = "Year", y = "Live Weight in '000 tonnes", fill="") + theme(plot.title = element_text(size=12, face="bold")) + facet_grid(.~Species.Group)  

        ggplotly(p2) %>% layout(margin=list(l=70, b=120)) 
       
                }  else if (string() == "Total") {
                    p2 <- ggplot(fisheries_priceplot2, aes(x=Year,y = Value/1000, fill=Type)) + geom_bar(position="dodge",stat="identity") + ggtitle("Total fish value over the years by Species Group") + labs(x = "Year", y = "Total Value in Million GBP", fill="") + theme(plot.title = element_text(size=12, face="bold")) + facet_grid(.~Species.Group)       
                          ggplotly(p2)%>%layout(margin=list(l=70, b=120))
                    
                } else {
                        
          p2 <- ggplot(fisheries_priceplot2, aes(x=Year,y=Price, fill=Type)) + geom_bar(position="dodge",stat="identity") + ggtitle("Average fish price over the years by Species Group") + labs(x = "Year", y = "Average Fish Price (GBP per tonne)", fill="") + theme(plot.title = element_text(size=12, face="bold")) + facet_grid(.~Species.Group) 
                        
 ggplotly(p2)%>%layout(margin=list(l=70, b=120))
      }
        })
        
# Key fishes landed by UK vessels in terms of quantity, total value and average price
    
        output$plot3 <- renderPlotly({
                 if (string() == "Live ") {
p3 <- ggplot(subset_species()[c(1:12),c(1:4)], aes(x = Year , y = Live.Weight/1000, fill=Species)) + geom_bar(position="dodge",stat = "identity") + labs(x = "Year", y = "Live Weight in '000 tonnes", fill="Species") + theme(plot.title = element_text(size=12, face="bold"))

ggplotly(p3) %>% layout(margin=list(l=70, b=180))
                         }
                else if (string() == "Total") {
p3 <- ggplot(subset_species()[c(13:24),c(1:3,5)], aes(x = Year, y = Value/1000, fill=Species)) + geom_bar(position="dodge",stat = "identity") + labs(x = "Year", y = "Total Value in Million GBP", fill="Species") + theme(plot.title = element_text(size=12, face="bold")) 

ggplotly(p3) %>% layout(margin=list(l=70, b=180))
                }
                else{
p3 <-  ggplot(subset_species()[c(25:36),c(1:3,6)], aes(x = Year, y = Price, fill = Species)) + geom_bar(position="dodge", stat = "identity") + labs(x = "Year", y = "Average Fish Price (GBP per tonne)", fill="Species") + theme(plot.title = element_text(size=12, face="bold"))  

ggplotly(p3) %>% layout(margin=list(l=70, b=180))
                }
        })
        
```

### Fisheries over months of the year
 
Please select criteria to view the fishing quality, total value and average price changing over 12 months for different years. <br/> Note: data for 2017 are up to June. 

```{r, context="render", echo=FALSE}

plotlyOutput("plot1")

```

### Fisheries by UK/foreign vessels

There are 3 species groups. 

<b>Demersal</b>: species living on or near the sea bed, such as cod, haddock, or plaice. <br/>
<b>Pelagic</b>: species found mainly in shoals in midwater or near the surface of the sea, such as herring or mackerel.<br/>
<b>Shellfish</b>: all crustaceans and molluscs, such as scallops, lobsters, or crabs. <br/>

```{r, context="render", echo=FALSE}

plotlyOutput("plot2")

```

### Key fishes landed by UK vessels

Here we are considering UK vessels' fisheries activity, excluding foreign vessels. Please select a species group to show top 3 fish caught in terms of live weight, total value and average price.  

```{r, context="render", echo=FALSE}

radioButtons("Species.Group", h4("Select a species:"),   inline = TRUE, choiceNames =c("Demersal", "Pelagic", "Shellfish"), choiceValues = c("Demersal", "Pelagic", "Shellfish"))

```

```{r, context ="server"}

subset_species <- reactive({
               Top3Species_plot3[Top3Species_plot3$Species.Group == input$Species.Group,]
        
})
```

```{r, context="render", echo=FALSE}

plotlyOutput("plot3")

```

Map of Ports {.page data-orientation=rows}
===================================== 

Here you can view key fisheries statistics for each year in the Data Viewer and navigate through the map for more information. 

Note: data for 2017 are up to June. 

### Data Viewer

```{r, context ="render", echo=FALSE}

selectInput("Year", label = h4("Select the Year"), choices=list("2017" = "2017","2016" = "2016","2015" = "2015", "2014" = "2014"), selected = "2017")

dataTableOutput("dataviewerperyear")

```

### Map 

```{r, context ="server"}

subset_year <- reactive({
              fisheriesData1[fisheriesData1$Year == input$Year,]
        
})

output$dataviewerperyear <- renderDataTable({
        subset_year() %>% 
group_by(Year, Port.of.Landing, Port.Nationality) %>% summarise(Live.Weight = sum(Live.weight..tonnes.),Value=sum(Value..000s.*1000))
        
})

output$map <- renderLeaflet({
                
                mapdata <- subset_year() %>% group_by(Port.of.Landing, Lon, Lat) %>% summarise(Live.Weight=sum(Live.weight..tonnes.),Value=sum(Value..000s.*1000))
                
leaflet(mapdata) %>% 
                        addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
                        addMarkers(clusterOptions=markerClusterOptions(),~Lon,~Lat,popup= ~htmlEscape(paste(Port.of.Landing, " - ", paste0("Year: ", input$Year), "  ", paste0("Quantity: ", Live.Weight," tonnes"), "  ", paste0("Worth: GBP ", Value))))
        })

```

```{r, context ="render", echo=FALSE}
leafletOutput("map", height=800)
```


